\documentclass[12pt,a4paper]{article}
%\usepackage{geometry}[paper=a4paper,left=30mm,width=150mm,top=25mm,bottom=25mm]
\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\author{Timothy Trewartha \\
\and
Michiel Johan Baird \\
\and
Supervisor: Hussein Suleman \\
Department of Computer Science \\
University of Cape Town
 }
\title{Spatial Navigation of African Cultural Heritage Sites}
\maketitle

\section{Project Description}
The Zamani project, started by the UCT Department of Geomatics, aims to
preserve African cultural heritage by documenting heritage sites and producing
laser scanned models. Currently they have documented about 40 sites in 12 African
countries with close to 100 models. Some of the models are very detailed containg
billions of points. With a fast growing volume of data the Department of Geomatics
is facing several challenges ranging from basic storage of the data to viewing and
interacting with the large models in real-time. This project aims to investigate
and develop solutions to these problems. Additionally the project will investigate
various ways of automating GIS workflow with a view to developing a software tool
to enable more efficient manipulation of GIS data.

\section{Problem Statement and Research Questions}
\subsection{Is it feasible to support real time viewing of models containing billions of points?}
The UCT Department of Geomatics has indicated that they have difficulties
handling the size of some of their models. These laser scanned models of
cultural heritage sites are often very large, some of them containing over
8 billion points. Given this vast scale of data, traditional viewing methods
and the current hardware and software systems are not able to cope.
Consequently, before viewing or manipulating the data, one must go through
a process of decimating the original data by a factor of 10, 100 or more.
This compromise is often unacceptable, as one may often require the full
original detail. This level of detail is often necessary for cultural heritage
sites in order to view details such as cracks and flaws, with a view to
preserving the site and preventing damage.
\subsection{Can we provide a central storage repository for the data without the server becoming a bottleneck?}
The Geomatics Department also mentioned that they have no central
storage location for their many models. Some models are stored on
a server, but the server does not have sufficient storage capacity
for all thier models. Some models are stored on client machines so
that the data is immediately available to be viewed and manipulated
as required. Lastly, there are some models that are simply stored on
external hard drives if there is no space on the server, and if they
are not currently being used. This leads to many issues concerning
data consistency, data availability and data safety. Ideally one should
have a central storage server, but this could introduce a significant
bottleneck to data access.
\subsection{Can GIS research be implemented with an automated workflow system?}
Geographic Information Science involves the capture, storage, manipulation and analysis
and management of geographic data. This data is very diverse and as such has to be handled
in quite diverse ways. This data gets abstracted into various forms. This presents a
rather unique challenge in managing the data as it could be used by anyone of the research
staff at any point in the process. This data movement is laborious and could benefit from
from automisation.

Workflow Management Systems aim to decompose compilated projects and processes into
small atomic chucks. This decompisition can then be optimised to improve the efficiency.
GIS research projects are generally done with multiperson teams where the work is
done in a parallel fashion. Under these conditions workflow management systems
are optimal.

The aim is to provide a workflow management that is applicable for GIS projects.
This system should be able to: interface with the current systems; track and
manage the workflow; provide local data availabilty and content delivery; and
increases overall efficiency within the dicipline.

\section{Procedures and Methods}
\subsection{Hierarchical Data Structure}
From researching the literature it seems that the most common way of dealing
with large point based models containing billions of points is to build a
multiresolution datastructure to divide our model into manageable chunks.
Initially, we need only a small subset of the number of available points.
As we zoom into the model we will need to request more points from the
datastructure until the full original detail is available. Using such a
level-of-detail structure should enable the Department of Geomatics to
view even very large models at interactive frame rates, without having
to decimate the original data.
\subsection{Server with 20TB storage capacity}
In order to have a central repository for all the models, we aim to obtain a server
with around 20TB of storage capacity. This will enable us to keep all the models in
a central location.
\subsection{Testing and Evaluation}
The most important evaluation criteria will be to demonstrate that the new
system is able to render the large point based models in realtime. Since this
functionality was not able previously, it will be a significant success.
Additionally it will be important to test whether realtime streaming from the
server to client machines is feasible.

\section{Ethical, Professional and Legal Issues}
\subsection{User Testing}
One of the componenets of the project invloves the design and evaluation
of a user interface. Ideally the design process for this would require
input from the user as well as testing. This testing requires ethical clearance
that will have to be obtained.
\subsection{Data Privacy}
The Department of Geomatics has indicated that some of the data collected by
the Zamani Project is sensitive and is not to be made freely available. It is
important to ensure that during the course of this project this wish is respected
and that nothing is done to compromise the privacy of sensitive data.



\section{Related Work}
\subsection{Hierarchical Data Structure}
Common methods for structuring 3D data include octrees \cite{interactivepointclouds},
R-trees \cite{rtree}, bounding sphere hierarchies \cite{qsplat}, and Hilbert Space
Filling Curves \cite{hilbert}, each with their own advantages and disadvantages. Based
on the experimental results of each method, a dynamic octree structure seems to have
the best performance \cite{interactivepointclouds}. Using this datastructure the
authors were able to achieve interactive walkthroughs of a data set with 2.2
billion points totalling 63.5GB.

\section{Anticipated Outcomes}

\section{Project Plan}
\subsection{Risks}
\subsubsection*{Request for Hardware Denied}
\noindent \textit{Severity: } High \\
\noindent \textit{Likelihood: } Medium \\
It is possible that the request for a server will be denied. If this happens
a considerable amount of restructuring will be required and it would have a
significant impact on the course of the project.
\subsubsection*{Network Constraints}
\noindent \textit{Severity: } Medium \\
\noindent \textit{Likelihood: } Low \\
One of the core functionalities of the system would be to provide
content delivery of data that is required for a specific task. Providing
this local data allows the task to get completed without unnessesry
fetching delays. There is a risk that this content delivery system
could saturate the network. This would cause the system to be slow
and unusable.
\subsubsection*{Middleware}
\noindent \textit{Severity: } High \\
\noindent \textit{Likelihood: } Low \\
For this project to be successful, the WFMS it would have to interface
heavily with existing software used to perform GIS operations. This will
require large amounts of middleware to be developed that understand the
the input and output formats of this software. Since many of these
formats are propriatary a significant amount of effort will have to
be made for the sytem to function. If these formats can not be intergrated
it presents a huge risk to the project.
\subsubsection*{Hardware Limitations}
\noindent \textit{Severity: } Medium \\
\noindent \textit{Likelihood: } Medium \\
There is a risk that the hardware available will not be able to cope
with the load that will be required. Since a distributed system is
not being proposed there is a risk that the system will become a bottleneck.
\subsubsection*{Large Indices}
\noindent \textit{Severity: } Medium \\
\noindent \textit{Likelihood: } Low \\
When indexing the models a significant amount of data will be generated.
Given that many of the models are already very large, these indices might
become infeasibly large. Dealing with such large indices will be an important
part of the project and this risk will have to be handled carefully.
\subsubsection*{Integration with Existing GIS Software}
\noindent \textit{Severity: } Low \\
\noindent \textit{Likelihood: } Medium \\
The hierarchical datastructure required as part of this project aims to facilitate level
of detail streaming and realtime interaction. However, ideally one should not have to
re-implement tools which are already available such as ArcGIS. The aim is to integrate
the datastructure into a pre-existing software package to prevent unnecessary work.
However, this may be difficult and there are several associated risks such as,
unavailability of source code, lack of documenation for the software, and potential
copyright license infringement.
\subsubsection*{Indexing takes too long}
\noindent \textit{Severity: } High \\
\noindent \textit{Likelihood: } Low \\
In the literature it had been noted that indexing the 3D data requires a significant
amount of time \cite{interactivepointclouds}. While it is hoped that this will not be
a problem, steps will have to be taken to ensure that the duration of the indexing
process does not pose a risk to the completion of our project. It will also be important
to allow time for the possible event of a systems failure, in which case the indexing
process would have to be restarted.
\subsection{Timeline, including Gantt chart}
We need a nice program for creating Gantt charts
\subsection{Resources required}
\subsubsection*{Hardware}
A server will be required for this project to enable central storage of the 3D models.
A large amount of storage will be required, as the Zamani data is over 20TB. Multiple
harddrives will be required and this will allow for a certain amount of parallelism in
data access. Additionally the server will ideally have the processing capacity to index
the models in a reasonable amount of time.
\subsubsection*{Geographic Data}
The Department of Geomatics has indicated that they are willing to make their data
available for the purposes of this project. It will be important to obtain the
models at an early stage as any delays in obtaining the models will delay the entire project.
\subsection{Deliverables}
\subsubsection*{GIS Workbench}
The core of the project is to produce a GIS workbench. This will
be the framework that ties all the components together. This will involve
using an existing Workflow System and setting it up to represent the
flow of a GIS project.
\subsubsection*{Middleware for Core Functionalities}
Once the GIS workflow is properly understood and modelled, it is
important to create middleware that interfaces with the systems that are
currently being used.
\subsubsection*{Data Flow Facilitator}
To be able produce the content delivery that is required, a dataflow
facilitator will need to be developed that is integrated with the
workflow management system.
\subsubsection*{Hierarchical Data Structure}
In order to facilitate level of detail streaming it will be essential to
implement a hierarchical data structure which can support interactive
viewing of models containing billions of points.
\subsubsection*{Streaming Infrastructure}
A realtime streaming infrastrucutre from the server to client machines will be
an important deliverable. This will also need to be implemented as early as
possible.

\subsection{Milestones}
\subsubsection{Project Milestones}

\begin{tabular}{l||l|l|l}
    Description & Duration & Start & Finish \\
    \hline  \hline
    Literature Review & 18 days & 24 April 2012 & 14 May 2012 \\
    \indent Fist Draft Due & 15 days  & 24 April 2012  & 9 May 2012  \\
    \indent Second Draft & 3 days & 9 May 2012 & 12 May 2012 \\
    \hline
    Project Proposal & 12 days & 10 May 2012  & 21 May 2012 \\
    \indent First Draft Due & 7 days & 10 May 2012 & 17 May 2012 \\
    \indent Second Draft Due & 2 days  & 17 May 2012  & 19 May 2012 \\
    \hline
    Project Proposal Presentation & 6 days  & 18 May 2012 & 24 May 2012 \\
    \indent Proposal Dry run & 4 days & 18 May 2012 & 22 May 2012 \\
    \hline
    Proposal Revision Completed & 7 days & 4 June 2012 & 11 June 2012 \\
    \hline
    Project Web Presence & 18 days  & 25 May 2012 & 12 June 2012 \\
    \hline
    Initial Feasibility Demonstration & 48 days & 11 June 2012 & 29 July 2012 \\
    \indent Design & 10 days & 11 June 2012 & 21 June 2012 \\
    \indent Implementation & 27 days & 21 June 2012 & 18 July 2012 \\
    \indent Evaluation & 11 days & 18 July 2012 & 29 July 2012 \\
    \hline
    Background Chapter & 28 days & 1 July 2012  & 29 July 2012 \\
    \indent First Draft Due & 21 days & 1 July 2012 & 22 July 2012 \\
    \indent Second Draft Due & 4 days  & 22 July 2012 & 26 July 2012 \\
    \hline
    Design Chapter & 31 Days & 29 July 2012 & 29 August 2012 \\
    \indent First Draft Due & 17 days  & 29 July 2012 & 15 August 2012 \\
    \indent Second Draft Due & 14 days & 15 August 2012  & 26 August 2012 \\
    \hline
    First implementation and Writeup & 31 days & 1 July 2012 & 29 August 2012 \\
    \indent Design & 13 days & 1 July 2012 & 14 July 2012 \\
    \indent Implementation & 31 days & 14 July 2012 & 15 August 2012 \\
    \indent Evaluation & 14 days & 15 August 2012 & 26 August 2012  \\
    \indent Writeup & 29 days & 30 July 2012 & 29 August 2012 \\
    \hline
    Final implementation and Writeup & 30 days & 29 August 2012 & 28 September 2012 \\
    \indent Design & 9 days   & 29 August 2012  & 7 September 2012 \\
    \indent Implementation & 14 days & 7 Semptember 2012 & 21 September 2012 \\
    \indent Evaluation & 7 days & 21 September 2012  & 28 September 2012 \\
    \indent Writeup & 30 days & 29 August 2012 & 28 September 2012 \\
    \hline
    Report Outline Complete & 12 days  &28 September 2012 & 10 October 2012 \\
    \hline
    Final Complete Draft of Report & 14 days & 10 October 2012 & 24 October 2012 \\
    \indent First Draft Due & 4 days &10 October 2012 & 14 October 2012 \\
    \indent Second Draft Due & 4 days &14 October 2012 & 18 October 2012 \\
    \indent Third Draft Due & 4 days  &18 October 2012 & 22 October 2012 \\
    \hline
    Project Report Final Handin & 7 days & 24 October 2012 & 31 October 2012 \\
    \hline
    Poster Due & 4 days  & 31 October 2012 & 3 November 2012 \\
    \hline
    Webpage & 4 days & 31 October 2012 & 3 November 2012 \\
    \hline
    Project Demonstrations & 5 days & 3 November 2012 & 8 November 2012 \\
    \hline
    Write Reflection paper &3 days & 8 November 2012 & 11 Noveber 2012 \\
    \hline
    Final project presentation & 7 days &11 November 2012 & 18 Noveber 2012 \\

\end{tabular}
\subsubsection{Implementation Milestones}

\subsubsection*{Datastructure Implemented}
It will be important to implement the datastructure at an early stage as it
forms the basis for several other tasks.
\subsubsection*{System is able to render large models}
Once the datastructure has been implemented it will be used to test the
rendering of one of the large models supplied by the Geomatics Department.
\subsubsection*{System is able to stream large models from the server}
Once the system is capable of rendering large models the next important
milestone will be to allow streaming of the model from the central server to
the client workstation.
\subsubsection*{Developed a model for basic GIS workflow}
Understanding how GIS research is vitally important for this part of
the project. This will allow a model to be developed. This model will
be the center of the design of the system. It will provide the
information required to pick an appropriate sytem.
\subsubsection*{Basic Middleware Implemented}
This will act as the link between the current software and the WFMS.
\subsubsection*{Content Delivery System}
This system will facilitate automated data transfer between the workstations
where it is required.
\subsubsection*{Implemented a User Interface}
The user interface will be the main control mechanism of the sytem. This
will allow the workflow to be monitored and adjusted according to the users
requirements.

\subsection{Work Allocation}
Timothy Trewartha will be implementing the hierarchical data structure to support
real-time interaction with the Zamani models. Michiel Johan Baird will be developing
the GIS workbench. A certain amount of integration between the two will also be required.




\bibliographystyle{apalike}
\bibliography{bibliography.bib}
\end{document}
